---
title: "TTstar runtime"
output: html_notebook
---

```{r,hidden}
# libraries & setup

library(ggplot2)
library(reshape2)
library(repr)

options(repr.plot.width=4, repr.plot.height=3)
```


```{r}
# load measured data
xs <- read.csv('benchmark-1490607733.csv')
#xs <- read.csv('benchmark-1491380156.csv')
str(xs)
```

```{r}
# convert strings "True"/"False" into proper R booleans
bools <- c('inference', 'specialisation', 'verification', 'normalisation', 'compilation')
nas <- xs[,bools] == ""
xs[,bools] <- xs[,bools] == "True"
xs[,bools][nas] <- NA
```

Next, we build configuration descriptors in the form `$stage-$features`.

There are three stages:

* `ttstar` -- takes TTstar source, emits Scheme source
* `csc` -- compiles Scheme source to native code (via GCC); not present if intepreting
* `execute` -- running the final program
  * either compiled in the `csc` stage
  * or interpreted using `csi` directly from the Scheme source
  
There are features that can be enabled or disabled:

* `i`-nference -- erasure inference + erasure
* `s`-pecialisation -- erasure polymorphism
* `v`-erification -- final type check of the fully explicitly erasure-annotated program
* `n`-ormalisation -- NF-reduction of the final program
* `c`-ompilation -- running `csc`, the Scheme compiler

If the flag `c` is not present, the Scheme program is interpreted using `csi`.

```{r}
# make config strings
tr <- xs

block <- xs[,bools]
block[is.na(block)] <- F

tr$config <- paste(as.character(tr$stage), "-", sep="")
for (col in bools) {
    tr$config[block[,col]] <- paste(tr$config[block[,col]], substr(col, 1, 1), sep="")
}
tr$config <- factor(tr$config)
```

## Stage: TTstar

We can clearly see that while normalisation and verification take just a little bit of time, erasure inference takes the bulk of processing time.
```{r}
summary(tr$config)
```

```{r}
# ttstar
ttstar <- tr[tr$stage == "ttstar",]

options(repr.plot.width=6, repr.plot.height=3)
ggplot(ttstar, aes(x=config, y=runtime)) +
    geom_boxplot() +
    coord_flip() +
    expand_limits(y=0) +
    ggtitle("TTstar runtime")
```

If we run multivariate linear regression, we learn that in this program:

* inference takes **100 ms**
* verification takes **2.2 ms**
* normalisation takes **1.6 ms**
* other overheads (parsing, testing harness, ...) take **9.3 ms**

```{r}
m <- lm(runtime ~ inference + verification + normalisation, data=ttstar)
summary(m)
```

## Stage: Chicken Scheme compiler

Here we can see that while erasure inference takes **100 ms** itself, it shaves **2 seconds** off code generation.

```{r}
# ttstar
csc <- tr[tr$stage == "csc",]

options(repr.plot.width=6, repr.plot.height=3)
ggplot(csc, aes(x=config, y=runtime)) +
    geom_boxplot() +
    coord_flip() +
    expand_limits(y=0) +
    ggtitle("Chicken Scheme compiler runtime")
```

```{r}
m <- lm(runtime ~ inference + verification + normalisation, data=csc)
summary(m)
```

```{r}
# ttstar
exec <- tr[tr$stage == "execution" & !tr$normalisation & !tr$verification,]

options(repr.plot.width=7, repr.plot.height=5)
ggplot(exec[exec$inference,], aes(x=input_size, y=runtime, color=config)) +
    geom_point(alpha=0.3) +
    ggtitle("TTstar runtime")
```

```{r}
ggplot(exec[!exec$compilation,], aes(x=input_size, y=runtime, color=config)) +
    geom_point(alpha=0.3) +
    ggtitle("TTstar runtime")
```

```{r}
ggplot(exec[!exec$inference,], aes(x=input_size, y=runtime, color=config)) +
    geom_point(alpha=0.3) +
    ggtitle("TTstar runtime")
ggsave("unerased-runtime.pdf", width=3.3, height=2.6, units='in')
```
